---
title: "FML_5"
author: "Samuel James Gudise"
date: "2024-04-07"
output: html_document
---

## Interpretation - 
The Euclidean distance metric is used in the hierarchical clustering process using standardized measurements. Agnes algorithm is then used to compare several linkage techniques, such as single, complete, average, and Ward. The best clustering strategy is chosen based on the obtained results. Data is partitioned and cluster assignments are compared between partitions to evaluate stability. Normalizing the data is crucial to identifying a cluster of "healthy cereals" appropriate for elementary school cafeterias, guaranteeing that all cereals contribute to a well-rounded and nutrient-dense diet for pupils.

Synopsis and conclusions:

It has been determined that four clusters is the ideal quantity.

Organizational Structure:

Cluster 1: Consists of cereals with a high nutritional content that are balanced in terms of fiber, protein, and low levels of sugar and fat.

Cluster 2: Contains cereals with a well-balanced mix of nutrients and mild nutritional characteristics.

Cluster 3: Have lesser fiber and a higher sugar content, making them possibly less nutrient-dense.

Cluster 4: Contains grains that are richer in sugar and fat content but have a lower overall nutritional quality.

Stability analysis confirms the validity of the four clusters, which identify different nutritional groups based on different data subsets.
Data normalization is essential for accurate cluster comparison. Cluster group 1, comprising (69, 61, 27, 44, 58, 64, 2, 65, 66, 55, 58, 21, 35, 18, 31, and 67).



```{r}
#loading the library 
library(dplyr)
library(cluster)
```

# Load the dataset
```{r}
Cereals_Data <- read.csv("./Cereals.csv")
```

#reamoving the rows which are missing values.
```{r}
Cere_clene <- na.omit(Cereals_Data)
head(Cere_clene)
```

```{r}
Num_columns <- sapply(Cereals_Data, is.numeric)
Cereal_Norm <- Cere_clene
Cere_scale <- Cereals_Data
Cere_scale[, Num_columns] <- scale(Cere_scale[, Num_columns])
head(Cere_scale)
```

```{r}
# Dissimilarity matrix
Diss_mat <- dist(Cereals_Data, method = "euclidean")
# Hierarchical clustering using Complete Linkage
Hi_clus<- hclust(Diss_mat, method = "complete")
# Plot the obtained dendrogram
plot(Hi_clus, cex = 0.6, hang = -1)

```

```{r}
# Euclidean distance matrix
Distance_mat <- dist(Cereal_Norm, method = "euclidean")

# Single Linkage
HClus_single <- hclust(Distance_mat, method = "single")

# Complete Linkage
HClus_complete <- hclust(Distance_mat, method = "complete")

# Average Linkage
HClus_average <- hclust(Distance_mat, method = "average")

# Ward's Method
HClus_Ward <- hclust(Distance_mat, method = "ward.D2")
par(mfrow = c(2,2))
plot(HClus_single, main = "Single Linkage", sub = "", xlab = "", ylab = "Distance")
plot(HClus_complete, main = "Complete Linkage", sub = "", xlab = "", ylab = "Distance")
plot(HClus_average, main = "Average Linkage", sub = "", xlab = "", ylab = "Distance")
plot(HClus_Ward, main = "Ward's Method", sub = "", xlab = "", ylab = "Distance")

```

```{r}
# Calculating the distance of Euclidean
Distance_EUC <- dist(Cereal_Norm, method = "euclidean")

# Implement hierarchical clustering with several linking techniques
HClus_single <- hclust(Distance_EUC, method = "single")
HClus_complete <- hclust(Distance_EUC, method = "complete")
HClus_average <- hclust(Distance_EUC, method = "average")
HClus_Ward <- hclust(Distance_EUC, method = "ward.D2")
par(mfrow=c(2,2))
plot(HClus_single, main = "Single Linkage", sub = "", xlab = "", ylab = "Distance")
plot(HClus_complete, main = "Complete Linkage", sub = "", xlab = "", ylab = "Distance")
plot(HClus_average, main = "Average Linkage", sub = "", xlab = "", ylab = "Distance")
plot(HClus_Ward, main = "Ward's Method", sub = "", xlab = "", ylab = "Distance")



```

```{r}
clusters <- cutree(HClus_Ward, k=5)
Cere_clene$Cluster <- clusters

# Exploring the clustering composition
table(Cere_clene$Cluster)

# Analyzing the mean values by the cluster
aggregate(Cere_clene[, -1], by = list(Cere_clene$Cluster), mean)

```

```{r}
#Example: Determine which cluster has the lowest average sugar content and the highest average fiber content.
#In order to modify appropriately, this phase may require domain knowledge.

Best_index <- with(aggregate(Cere_clene[, c("fiber","sugars")], by = list(Cere_clene$Cluster),mean), which.max(fiber - sugars))
Best_Cereals <- Cere_clene[Cere_clene$Cluster == Best_index, ]
```

```{r}
library(cluster)
Data_Fcer <- Cereals_Data
#Compute using agnes and various linking techniques
HClus_single <- agnes(Cereals_Data, method = "single")
HClus_complete <- agnes(Cereals_Data, method = "complete")
HClus_average <- agnes(Cereals_Data, method = "average")

```

```{r}
print(HClus_single$ac)
print(HClus_complete$ac)
print(HClus_average$ac)
```

```{r}
pltree(HClus_complete, cex = 0.6, hang = -1, main = "dendrogram of agnes")
```

```{r}
Data_Fcer <- Cereals_Data
dat <- dist(Cereals_Data, method = "euclidean")
```

```{r}
HClus_complete <- hclust(dat, method = "complete")
```

```{r}
plot(HClus_complete, cex = 0.6)
rect.hclust(HClus_complete, k = 4, border = 1:4)
```

```{r}
HCLus_Diana <- diana(Cereals_Data)
HCLus_Diana$dc
```

```{r}
pltree(HCLus_Diana, cex = 0.6, hang = -1, main = "dendrogram of diana")
```

